---
title: "NVD stock 2018-2024"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    code_folding: show
---

## data prep
```{r}
library(dplyr)
library(readr)

# Read CSV
nvidia <- read_csv("data/NVIDIA_STOCK.csv", col_names = FALSE)

# Use FIRST ROW as proper column names (Price, Adj Close, Close...)
colnames(nvidia) <- as.character(unlist(nvidia[1, ]))

# Remove the first THREE rows
df <- nvidia[-c(1, 2, 3), ]

# Rename the first column "Price" → "Date"
df <- df %>%
  rename(Date = Price)

# Convert types
df <- df %>%
  mutate(
    Date = as.Date(Date),
    `Adj Close` = as.numeric(`Adj Close`),
    Close = as.numeric(Close),
    High = as.numeric(High),
    Low = as.numeric(Low),
    Open = as.numeric(Open),
    Volume = as.numeric(Volume)
  )

# Check
tail(df)
str(df)
# write the clean data to a new .csv file
write_csv(df, "NVIDIA_STOCK_CLEAN.csv")
```

## Close Price Over Time
```{r}
library(ggplot2)

ggplot(df, aes(x = Date, y = `Adj Close`)) +
  geom_line(color = "#0073C2FF", size = 1) +
  labs(title = "NVIDIA Adjusted Close Price Over Time",
       x = "Date",
       y = "AdjClose Price (USD)")

```

1. Close Price Over Time

Interpretation

NVIDIA's stock price demonstrates remarkable growth over the 6-year period, transforming from approximately $5-10 in 2018 to peaks exceeding $130 by late 2024. The trajectory reveals four distinct phases:

- Phase 1 (2018-2020): Steady Foundation - The stock traded relatively flat between $5-15, showing modest growth with minimal volatility. This period reflects NVIDIA's transition from primarily a gaming GPU company to diversifying into data centers.

- Phase 2 (2020-2021): Cryptocurrency Surge - A sharp rally from ~$10 to $35 occurred, driven by the cryptocurrency mining boom which created massive demand for NVIDIA's GPUs. The stock peaked around $35 in late 2021.

- Phase 3 (2022-early 2023): Correction Period - A significant pullback to ~$10-15 reflects the crypto market crash, global chip shortage concerns, and broader market correction during the Federal Reserve's interest rate hiking cycle.

- Phase 4 (2023-2024): AI Revolution Explosion - The most dramatic growth phase, where the stock surged from $15 to over $130 (approximately 8-9x increase). This unprecedented rally was fueled by the generative AI boom following ChatGPT's launch.

Key Events Correlation

2018-2019
- September 2018: RTX 20-series launch (first ray-tracing GPUs) - modest price impact
- Crypto mining bubble burst led to GPU oversupply and price stagnation

2020
- May 2020: Mellanox acquisition completed ($7 billion) - strengthened data center position
- September 2020: RTX 30-series "Ampere" launch - significant performance leap
- Stock began steady climb as data center revenue grew

2021
- Q1 2021: Data center revenue surpassed gaming for first time
September 2021: Grace CPU announcement - expansion beyond GPUs
Cryptocurrency mining demand drove stock to $35 peak
Stock split 4-for-1 in July 2021 (adjusted in data)

2022
- Crypto crash throughout 2022 caused sharp decline
- September 2022: RTX 40-series launch, but demand was soft
- US export restrictions on advanced AI chips to China announced (October)
- Stock hit bottom around $10-12

2023: The AI Transformation Year
- November 2022-March 2023: ChatGPT launches, sparking generative AI race
- May 2023: Blockbuster Q1 earnings - data center revenue doubled
- Stock began meteoric rise from $25 to $50+ by mid-2023
- Every major tech company (Microsoft, Google, Meta, Amazon) announced massive AI infrastructure investments using NVIDIA GPUs

2024
- March 2024: GTC Conference - announced Blackwell GPU architecture
- Stock reached all-time highs above $130
- NVIDIA briefly became the world's most valuable company (June 2024)
- 10-for-1 stock split in June 2024 (adjusted in data)
- Continued volatility but sustained high valuation driven by H100/H200 GPU demand


## Trading Volume
```{r}

ggplot(df, aes(x = Date, y = Volume)) +
  geom_col(fill = "#FF8C00", width = 1) +
  labs(
    title = "NVIDIA Trading Volume Over Time",
    x = "Date",
    y = "Volume (Shares Traded)"
  )


library(lubridate)
df_monthly <- df %>%
  group_by(YearMonth = floor_date(Date, "month")) %>%
  summarise(Volume = sum(Volume))

ggplot(df_monthly, aes(x = YearMonth, y = Volume)) +
  geom_col(fill = "#FF8C00") +
  labs(title = "NVIDIA Monthly Trading Volume",
       x = "Month",
       y = "Volume")

```

2. Monthly Trading Volume

Interpretation

Trading volume shows significant variability throughout the period, with clear spikes corresponding to major market events and earnings announcements. Volume patterns reveal investor sentiment and market activity intensity.

Key Volume Patterns:

- 2018: Moderate baseline volume (10-15 billion shares monthly)
- Early 2019: Extremely high volume spikes (~17 billion) during crypto bust uncertainty
- 2020: Elevated volume during COVID-19 market volatility and recovery
- 2021: High sustained volume during crypto bull run
- 2022: Reduced volume during bear market (5-10 billion range)
- 2023-2024: Moderate to high volume as AI narrative developed

Volume Event Correlation

High Volume Spikes Indicate:

- Earnings announcements (quarterly pattern visible)
- Major product launches
- Significant news events (acquisitions, restrictions, breakthroughs)
- Market regime changes (bull to bear transitions)

Notable Volume Events:

- Early 2019: Massive volume during crypto aftermath and gaming GPU glut
- March 2020: COVID-19 crash caused volume spike as investors repositioned
- May 2023: Record volume after stunning AI-driven earnings beat
- Mid-2024: Sustained high volume as institutional investors piled into AI trade

Volume typically spikes 1-2 days around earnings dates (typically late February, late May, late August, late November each year).

## Dynamic Candlestick Chart
```{r}
library(plotly)
plot_ly(data = df, x = ~Date, type = "candlestick",
        open = ~Open, close = ~Close,
        high = ~High, low = ~Low) %>%
  layout(title = "NVIDIA Candlestick Chart",
         xaxis = list(title = "Date"),
         yaxis = list(title = "Price (USD)"))
```

3. Candlestick Chart

Interpretation

The candlestick chart provides granular insight into daily price action, revealing buying/selling pressure through the relationship between opening, closing, high, and low prices.

Pattern Analysis:

- 2018-2020: Narrow Trading Range

Small candlestick bodies indicate low daily volatility
Prices consolidated as company transitioned business model
Red (down) and green (up) days relatively balanced

- 2021: Crypto Bull Run

Larger green candles show strong buying pressure
Upper wicks indicate resistance at higher levels
Steady uptrend with pullbacks

- 2022: Bear Market

Dominant red candles during downtrend
Long lower wicks show intraday selling pressure
Gap-down days following negative news

- 2023-2024: AI Boom

Massive green candles post-earnings (particularly May 2023)
Increased candle body sizes reflect higher daily price swings
Strong uptrend with occasional violent selloffs (volatility expansion)
Upper and lower wicks lengthened, showing increased intraday trading ranges

Technical Signals:

- Long green candles with small wicks = strong bullish momentum (2023-2024)
- Doji candles (small bodies, long wicks) = market indecision (visible in 2022)
- Gap ups after earnings = extremely positive market reaction (May 2023, several 2024 dates)


## Cumulative Return
```{r}
library(tidyr)
# Calculate the daily rate of return
# Return_t = Close_t / Close_(t-1) - 1
df <- df %>%
  arrange(Date) %>%
  mutate(Return = Close / lag(Close) - 1)

# Calculate the cumulative earnings
# Cumulative_Return_t = Π (1 + Return_i)
df <- df %>%
  mutate(Cumulative_Return = cumprod(1 + replace_na(Return, 0)))

# Plot
ggplot(df, aes(x = Date, y = Cumulative_Return)) +
  geom_line(color = "brown", size = 1) +
  labs(
    title = "NVIDIA Cumulative Return Over Time",
    x = "Date",
    y = "Cumulative Return (Growth Factor)"
  )

```

4. Cumulative Return Over Time

Interpretation

The cumulative return chart dramatically illustrates NVIDIA's extraordinary performance as an investment, with a growth factor exceeding 25x from 2018 baseline.
Return Analysis:

Growth Phases:

- 2018-2020: Cumulative return ~1-2x (minimal growth, basically flat)
- 2020-2021: Grew to ~7-8x (cryptocurrency and pandemic tech surge)
- 2022: Declined to ~4-5x (significant drawdown during bear market)
- 2023-2024: Exploded to ~25-27x (AI-driven growth)

Key Insights:

An investor who bought $10,000 of NVIDIA stock in early 2018 would have ~$250,000+ by late 2024
The 2022 drawdown of ~50% from peaks tested investor conviction
The recovery and new highs in 2023-2024 rewarded long-term holders
NVIDIA dramatically outperformed the S&P 500 (see comparison chart)

Risk-Adjusted Perspective:

While returns are spectacular, the journey included a 50%+ drawdown in 2022, highlighting the volatility risk investors faced. The crypto boom-bust cycle and subsequent AI boom show NVIDIA's dependency on emerging technology trends.


## 30-Day Rolling Volatility

Volatility: sd of the daily yield over the past N days
\[
\sigma_{30}(t) 
= \sqrt{\frac{1}{30} \sum_{i=0}^{29} \left( r_{t-i} - \bar{r}_{30}(t) \right)^2 }
\]

```{r}
library(zoo)

# Rolling 30-day Volatility）

df <- df %>%
  mutate(
    Volatility30 = rollapply(Return, width = 30, FUN = sd, fill = NA, align = "right")
  ) # "Return" is the daily rate of return from last part

# plot
ggplot(df, aes(x = Date, y = Volatility30)) +
  geom_line(color = "darkgreen", size = 1) +
  labs(
    title = "30-Day Rolling Volatility of NVIDIA",
    x = "Date",
    y = "Volatility (Std Dev of Daily Returns)"
  )

```

5. 30-Day Rolling Volatility

Interpretation

Volatility measures the standard deviation of daily returns, indicating price uncertainty and risk. Higher volatility periods correspond to uncertain market conditions or major news events.

Volatility Regime Analysis:

- 2020 COVID Crash: Highest volatility spike in entire dataset (~0.08) as markets faced unprecedented uncertainty
- 2022 Crypto Crash: Elevated volatility (~0.04-0.05) as NVIDIA navigated GPU oversupply and demand destruction
- 2023 AI Rally: Surprisingly moderate volatility given rapid price appreciation, suggesting strong institutional conviction

Investment Implications:

- High volatility periods offered entry opportunities for risk-tolerant investors
- Volatility tends to cluster (volatile periods followed by more volatility)
- Low volatility during 2023-2024 rally suggests strong conviction in AI narrative
- Options traders would have found rich premiums during 2020 and 2022 spikes


## Comparison with S&P500
```{r}
library(quantmod)

# Get S&P 500 data from Yahoo Finance (also the default source)
getSymbols("^GSPC", src = "yahoo", from = "2018-01-01", to = "2024-12-31")
head(GSPC)

# Convert to regular data frame
sp500 <- data.frame(Date = index(GSPC), coredata(GSPC))

# Rename columns to match your df style
colnames(sp500) <- c("Date", "Open", "High", "Low", "Close", "Volume", "Adjusted")

sp500 <- sp500 %>%
  arrange(Date) %>%
  mutate(Return = Close / lag(Close) - 1)

```

```{r}

sp500 <- sp500 %>%
  mutate(Cumulative_Return = cumprod(1 + ifelse(is.na(Return), 0, Return)))


# Merge by date
compare_df <- df %>%
  select(Date, NVDA_Close = Close, NVDA_Cum = Cumulative_Return) %>%
  inner_join(
    sp500 %>% select(Date, SP500_Close = Close, SP500_Cum = Cumulative_Return),
    by = "Date"
  )

# Plot cumulative return comparison
ggplot(compare_df, aes(x = Date)) +
  geom_line(aes(y = NVDA_Cum, color = "NVIDIA"), size = 1.1) +
  geom_line(aes(y = SP500_Cum, color = "S&P 500"), size = 1.1) +
  scale_color_manual(values = c("NVIDIA" = "#E74C3C", "S&P 500" = "skyblue")) +
  labs(title = "Cumulative Return: NVIDIA vs S&P 500",
       x = "Date",
       y = "Cumulative Return (Growth Factor)",
       color = "Legend")
```

6. Cumulative Return: NVIDIA vs S&P 500

Interpretation

This comparison chart starkly illustrates NVIDIA's massive outperformance versus the broader market, demonstrating both the company's exceptional growth and its higher risk profile.

Performance Comparison:
- NVIDIA: ~25x cumulative return (2,500% gain)
- S&P 500: ~2x cumulative return (~100% gain)
- Outperformance Ratio: NVIDIA returned approximately 12-13x more than the S&P 500 over this period.

Key Observations:

- 2018-2020: Both assets moved relatively similarly, with NVIDIA showing slightly more volatility but comparable returns. This period shows NVIDIA trading more like a "normal" tech stock.
- 2020-2021: NVIDIA began separating from S&P 500, driven by crypto demand and data center growth. The gap widened from 2x to ~3x outperformance.
- 2022 Correction: NVIDIA fell more sharply than S&P 500 (higher beta), demonstrating the downside of volatility. While S&P 500 declined modestly, NVIDIA gave back much of its gains, highlighting sector-specific risks (crypto crash, chip restrictions).
- 2023-2024 Divergence: The performance gap exploded dramatically. While S&P 500 gained steadily (~20-30% total), NVIDIA skyrocketed 400-500% during the same period. This reflects:

NVIDIA's unique positioning as the "picks and shovels" provider of the AI revolution
Virtually monopolistic position in AI training chips (90%+ market share)
Every major tech company competing to secure NVIDIA GPU supply

Risk-Return Profile:

- Beta Analysis: NVIDIA exhibits much higher beta (systematic risk) than market
- Correlation: Strong correlation during calm periods, divergence during sector-specific events
Diversification: NVIDIA's returns show it's not just a tech stock but a specialized AI infrastructure play

Investment Lesson:

While NVIDIA dramatically outperformed, investors had to endure:

- 50%+ drawdown in 2022
- Extreme day-to-day volatility
- Concentration risk (single stock vs. diversified index)
- Sector rotation risk (gaming → crypto → AI)

The S&P 500's steady, less volatile growth represents the value of diversification, even if absolute returns were much lower.